Abstract Data Type Inference

### 2015-03-16 algorithm.tex
Soft description of the formulas.

TODO DESCRIBE HOW TO GET THE FORMULA $F$ FROM $B$, AND ARGUE THAT WE HAVE
SOLVED THE PROBLEM.

@@

  “The symbolic ADT inference problem is computable for normal ADTs.”

>>>>

WRITE A CLEAN SEGUE TO THE NEXT PART

Let $h = \tup{O,<,c,f,m}$ be a history with operations $O = \set{ o_1, \ldots,
o_n }$. Without loss of generality, we assume the match targets $\set{o_1,
\ldots, o_k}$ are indexed consecutively from $1$ to $k$, for some $k < n$. We
define the macro $\textsc{Embed}_o(x,Y,z)$:
\begin{align*}
  & (c(o) \implies {\sf c}(x)) \land {\sf f}(x) = f(o) \\
  & \land (o \in \dom(m) \implies \lnot{\sf um}(x) \land {\sf m}(x) = z) \\
  & \land (o \not\in \dom(m) \implies {\sf um}(x)) \land \bigwedge_{y \in Y} x < y
\end{align*}
capturing the correspondence between the logical variable $x$ and operation
$o$. The variables $Y$ and $z$ represent the operations ordered after $x$, and
the operation which $x$ matches. The macro $\textsc{Identical}(x,y)$, defined as
\begin{align*}
  {\sf f}(x) = {\sf f}(y)
  \land {\sf c}(x) \Leftrightarrow {\sf c}(y)
  \land {\sf m}(x) = {\sf m}(y)
\end{align*}
captures whether the operations bound to $x$ and $y$ are identical.
To express the constraints among the matches of an embedded history, we define
the macro $\textsc{Match}(Y,z)$:
\begin{align*}
  \forall x.\ {\sf m}(x) = z \implies
  \textsc{ReadOnly}(x) \lor \bigvee_{y \in Y} \textsc{Identical}(x,y)
\end{align*}
which requires any operation which matches $z$ to be either read-only, or
identical to some operation in $Y$, which represents the operations of $h$
which match $z$. Finally, we express the embedding of $h$ with the macro
$\textsc{Embed}_h$:
\begin{align*}
  \exists x_1, \ldots, x_n.\ \bigwedge_{i=1}^{n} \textsc{Embed}_{o_i}(x_i,Y_i,z_i)
  \land \bigwedge_{i=1}^{k} \textsc{Match}(M_i, x_i)
\end{align*}
where $Y_i = \set{ x_j : o_i < o_j }$ are the variables corresponding to
operations ordered after $o_i$, and $z_i$ is the variable corresponding to
$m(o_i)$, if defined, and $M_i = \set{ x_j : m(o_j) = o_i }$ are the variables
corresponding to operations matching $o_i$.

  $h_1 \models \textsc{Embed}_{h_2}$ if{f} $h_1 \succeq h_2$.

We obtain a formula representing an ADT $A$ by taking the conjunction of the
negative embedding formulæ $\set{ \lnot\textsc{Embed}_h : h \in H }$ from any
set $H$ that generates the complement of $A$.

  The formula $\bigwedge_{h \in H} \neg \textsc{Embed}_h$ represents $A$,
  for any set $H$ which generates the compliment of $A$.

For normal and predictable ADTs, such sets are computable, and so our inference
problem is also computable.

@@

  The symbolic ADT inference problem is computable for normal and predictable ADTs.
